---
title: "MTHS3003_Coursework_20498015_Solutions"
author: "Akshay Bhandari"
date: "11 December 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
Coffee_Tuesday=read.csv("Coffee_Tuesday_20498015.csv")
Coffee_early=read.csv("Coffee_early_20498015.csv")
Coffee_weekend=read.csv("Coffee_weekend_20498015.csv")
Coffee_income=read.csv("Coffee_income_20498015.csv")
```

1. Provide an exploratory analysis of the number of customers served in 5 minute intervals using the data
provided in Coffee_Tuesday_20498015.csv .

```{r}
mean(Coffee_Tuesday$Number.served) #MEAN
max(Coffee_Tuesday$Number.served) #MAXIMUM NUMBER OF CUSTOMERS VISITING IN A 5 MINUTE INTERVAL
min(Coffee_Tuesday$Number.served) #MINIMUM NUMBER OF CUSTOMERS VISITING IN A 5 MINUTE INTERVAL
summary(Coffee_Tuesday$Number.served) #SUMMARY OF THE DATA INCLUDING THE INTERQUARTILE RANGES
var(Coffee_Tuesday$Number.served) #VARIANCE
sd(Coffee_Tuesday$Number.served) #STANDARD DEVIATION

timeslot1=Coffee_Tuesday$Number.served[1:24] #DATA DIVIDED FOR 8AM-10AM
timeslot2=Coffee_Tuesday$Number.served[25:48] #DATA DIVIDED FOR 10AM-12PM
timeslot3=Coffee_Tuesday$Number.served[49:72] #DATA DIVIDED FOR 12PM-2M
timeslot4=Coffee_Tuesday$Number.served[73:96] #DATA DIVIDED FOR 2PM-4PM
timeslot5=Coffee_Tuesday$Number.served[97:120] #DATA DIVIDED FOR 4PM-6PM

sumtimeslot=0
sumtimeslot[1]=sum(timeslot1)
sumtimeslot[2]=sum(timeslot2)
sumtimeslot[3]=sum(timeslot3)
sumtimeslot[4]=sum(timeslot4)
sumtimeslot[5]=sum(timeslot5)
sumtimeslot

meantimeslot=0
meantimeslot[1]=mean(timeslot1)
meantimeslot[2]=mean(timeslot2)
meantimeslot[3]=mean(timeslot3)
meantimeslot[4]=mean(timeslot4)
meantimeslot[5]=mean(timeslot5)

boxplot(timeslot1,timeslot2,timeslot3,timeslot4,timeslot5,xlab="Timings",ylab="No. of Customers",names=c("8am-10am","10am-12pm","12pm-2pm","2pm-4pm","4pm-6pm"),col=c("darkred"),border="blue")

plot(meantimeslot[1:5],bg='red',col='blue',lwd=3,cex=3,pch=21,ylab="Mean No of Customers")
text(1.2,6.5,"8am-10am")
text(2,5,"10am-12pm")
text(3,6,"12pm-2pm")
text(4,5,"2pm-4pm")
text(4.8,2.5,"4pm-6pm")
```

2. The coffee owner proposes that customer demand in a 5 minute interval follows a X ~ Po($\lambda$) distribution.
Estimate the customer demand during the weekday early morning (8:30-9am) busy period using
Coffee_early_20498015.csv. Due to staff and space constraints, the maximum number of customers
that can be served in a 5 minute interval is 12. 

a. Find the maximum likelihood estimate of $\lambda$.

```{r}
lambda=mean(Coffee_early$Number.served)
round(lambda,3) #OBSERVED MEAN FROM THE DATA
n=length(Coffee_early$Number.served)
n #TOTAL NUMBER OF OBSERVATIONS
t=sum(Coffee_early$Number.served) #TOTAL NUMBER OF CUSTOMERS VISITING THE COFFEE SHOP
lam = seq(6, 10, by=.01) #SEQUENCE TO FIND THE LAMBDA VALUE IN
like = dpois(t, n*lam) #LIKELIHOOD FUNCTION
plot(lam, like, type="l", lwd=2)
abline(v = lambda, col="red") #LINE AT MAXIMUM LIKELIHOOD VALUE
maxlam=lam[like==max(like)] 
maxlam #MLE
```

Note: Suppose that an extra member of staff would enable the coffee shop to serve an additional 3
customers per 5 minute interval if there is the demand

b. Estimate how often the extra member of staff will be busy.

```{r}
ppois(13,maxlam,lower.tail = FALSE,log=FALSE) #PROBABILITY OF GETTING 13 OR MORE CUSTOMERS IN A 5 MINUTE INTERVAL
##CONCLUDED IN THE REPORT
```

c. Given that an extra member of staff would cost 20 per hour and the average profit per customer
is 2.10, is it worth the owner having an extra member of staff between 8:30-9:00am?

```{r}
profitpercustomer= 3*2.1
costofextrastaff=20
profitearned=profitpercustomer-costofextrastaff/2
##EXPLAINED IN THE REPORT
```

3. The coffee shop owner suspects that the early morning busy period varies by day of the week.

a. Does the data in Coffee_early_20498015.csv support the coffee owners suspicions?

```{r}
earlymonday=Coffee_early$Number.served[Coffee_early$Day=='Monday']
earlytuesday=Coffee_early$Number.served[Coffee_early$Day=='Tuesday']
earlywednesday=Coffee_early$Number.served[Coffee_early$Day=='Wednesday']
earlythursday=Coffee_early$Number.served[Coffee_early$Day=='Thursday']
earlyfriday=Coffee_early$Number.served[Coffee_early$Day=='Friday']

meanearly=0
meanearly[1]=mean(earlymonday)
meanearly[2]=mean(earlytuesday)
meanearly[3]=mean(earlywednesday)
meanearly[4]=mean(earlythursday)
meanearly[5]=mean(earlyfriday)
meanearly #VECTOR WITH MEAN VALUES FOR EACH WEEKDAYS

mondaydata<-rep(NA,10) #VECTOR WITH 10 'NA' OBJECTS FOR EACH DAY
tuesdaydata<-rep(NA,10)
wednesdaydata<-rep(NA,10)
thursdaydata<-rep(NA,10)
fridaydata<-rep(NA,10)
for(i in 1:10){
  mondaydata[i]=sum(Coffee_early$Number.served[Coffee_early$Week == i & Coffee_early$Day == "Monday"]) 
  tuesdaydata[i]=sum(Coffee_early$Number.served[Coffee_early$Week == i & Coffee_early$Day == "Tuesday"])
  wednesdaydata[i]=sum(Coffee_early$Number.served[Coffee_early$Week == i & Coffee_early$Day == "Wednesday"])
  thursdaydata[i]=sum(Coffee_early$Number.served[Coffee_early$Week == i & Coffee_early$Day == "Thursday"])
  fridaydata[i]=sum(Coffee_early$Number.served[Coffee_early$Week == i & Coffee_early$Day == "Friday"])
} #THIS LOOP STORES SUM OF ALL CUSTOMERS ON SPECIFIC WEEKDAYS , PER WEEK BASIS

boxplot(mondaydata,tuesdaydata,wednesdaydata,thursdaydata,fridaydata,xlab="Days of week",ylab="Customers per week",names=c("Monday","Tuesday","Wednesday","Thursday","Friday"),col=c("darkred"),border="blue")

boxplot(earlymonday,earlytuesday,earlywednesday,earlythursday,earlyfriday,xlab="Days of week",ylab="Customers per day",names=c("Monday","Tuesday","Wednesday","Thursday","Friday"),col=c("darkred"),border="blue")
##CONCLUSIONS ABOUT THIS IS MADE IN THE RERPOT
```

b. How does the answer to question 3.(a) affect the conclusions drawn in question 2?

```{r}
#CONCLUSION GIVEN IN THE REPORT
```

4. The coffee shop owner is interested in weekend customer numbers. Use the data in Coffee_weekend_20498015.csv to address the following questions and discuss your conclusions. 

a. Are the data consistent with the coffee shop owners assumption that the mean number of
customers on Saturdays and Sundays are 250 and 230, respectively?

```{r}
data_saturday=Coffee_weekend$Saturday
t.test(data_saturday , mu=250)#TESTING MEAN OF SATURDAY AGAINST THE MEAN 250
data_sunday=Coffee_weekend$Sunday
t.test(data_sunday , mu=230)#TESTING MEAN OF SUNDAY AGAINST THE MEAN 230
```

b. Are the data consistent with the coffee shop owners assumption that the mean difference in the
number of customers between Saturday and Sunday is at least 20?

```{r}
t.test(data_saturday , data_sunday ,mu=20, alternative = "less")#TESTING MEAN DIFFERENCE BETWEEN SATURDAY AND SUNDAY
```


5. The coffee shop owner wants to know how income per day (response variable) depends on the following
covariates: 

. Day of the week

. Mean number of staff working throughout the day

. Number of meal options

. Daily temperature (oC)

. Term time (yes/no)

a. Find the best model for predicting income using the data in Coffee_income_20498015.csv.

```{r}
income=Coffee_income$Income #LOADING INCOME DATA
weekday=Coffee_income$Weekday #LOADING DAY OF THE WEEK DATA
nostaff=Coffee_income$NoStaff #LOADING NUMBER OF STAFF DATA
term=Coffee_income$Term #LOADING TERM DATA
mealopt=Coffee_income$MealOpt #LOADING MEAL OPTIONS DATA
temp=Coffee_income$Temp #LOADING TEMPERATURE DATA
model<-lm(income~ weekday +nostaff + term + mealopt + temp) #MODEL WITH ALL COVARIATES
summary(model)

AIC(model) #AIC OF THE FULL MODEL

bestmodel<-step(model)#STEP FUNCTION TO FIND THE BEST FIT MODEL
summary(bestmodel)

AIC(bestmodel)#AIC OF THE BEST FIT MODEL

```

b. Comment on the model fit and effect of the selected covariates on income.

```{r}
hist(bestmodel$residuals,xlab="Residuals",col = "darkred",main="",freq=FALSE)#HISTOGRAM OF RESIDUALS TO CHECK NORMLITY
curve(dnorm(x,0,sd(bestmodel$residuals)),-300,300,col=4,add=TRUE) 

plot(bestmodel$fitted.values,bestmodel$residuals,bg='red',col='blue',lwd=3,cex=3,pch=21,xlab="Fitted values",ylab="Residuals")
##SCATTER PLOT TO CHECK HOMOGENITY 
##COMMENTS MADE ARE IN THE REPORT
```

c. Estimate the level of income on a day with the following covariate information:

A Saturday in term time with daily temperature 15.9oC, on average 5 members of staff working
and 4 meal options.

```{r}
pred<-predict(model,newdata=data.frame(weekday='Saturday',term=1,temp=15.9,nostaff=5,mealopt=4))
pred #PREDICTING INCOME AT THE ABOVE COVARIATES VALUES FOR OUR FULL MODEL
predbestmodel<-predict(bestmodel,newdata=data.frame(weekday='Saturday',term=1,temp=15.9,nostaff=5,mealopt=4))
predbestmodel #PREDICTING INCOME AT THE ABOVE COVARIATES VALUES FOR OUR BEST MODEL

```